---
title: 'Bios 6301: Assignment 2'
author: "Lydia Yao"
output:
  pdf_document: default
  html_document: default
---

*Due Tuesday, 21 September, 1:00 PM*

50 points total.

Add your name as `author` to the file's metadata section.

Submit a single knitr file (named `homework2.rmd`) by email to michael.l.williams@vanderbilt.edu.
Place your R code in between the appropriate chunks for each question.
Check your output by using the `Knit HTML` button in RStudio.

1. **Working with data** In the `datasets` folder on the course GitHub repo, you will find a file called `cancer.csv`, which is a dataset in comma-separated values (csv) format. This is a large cancer incidence dataset that summarizes the incidence of different cancers for various subgroups. (18 points)

    1. Load the data set into R and make it a data frame called `cancer.df`. (2 points)
```{r}
cancer.df <- read.csv("cancer.csv")
```
   
    2. Determine the number of rows and columns in the data frame. (2)
```{r}
dim(cancer.df)
```

    3. Extract the names of the columns in `cancer.df`. (2)
```{r}
colnames(cancer.df)
```

    4. Report the value of the 3000th row in column 6. (2)
```{r}
cancer.df[3000,6]
```

    5. Report the contents of the 172nd row. (2)
```{r}
cancer.df[172,]
```

    6. Create a new column that is the incidence *rate* (per 100,000) for each row. The incidence rate is the `(number of cases)/(population at risk)`, which in this case means `(number of cases)/(population at risk) * 100,000`. (3)
```{r}
cancer.df["rate"] = (cancer.df$incidence/cancer.df$population)*100000
```

    7. How many subgroups (rows) have a zero incidence rate? (2)
```{r}
nrow(cancer.df[cancer.df$rate == 0,])
```

    8. Find the subgroup (rows) with the highest incidence rate.(3)
```{r}
cancer.df[which.max(cancer.df$rate),]
```
2. **Data types** (10 points)

    1. Create the following vector: `x <- c("5","12","7")`. Which of the following commands will produce an error message? For each command, Either explain why they should be errors, or explain the non-erroneous result. (4 points)

            max(x)
            sort(x)
            sum(x)
    `sum(x)` will produce an error message.  Because the list created is a list of characters, we are unable to sum them thus we get the error `invalid 'type' (character) of argument`.  `max(x)` will give the max of the character.  Since 7 is after both 5 and 1, it is our max.  `sort(x)` will sort the list based on the first characters thus give us `"12" "5" "7"`

    2. For the next two commands, either explain their results, or why they should produce errors. (3 points)

            y <- c("5",7,12)
            y[2] + y[3]
    Line two of the code above should return errors because of type incompatibility.  We can not sum two characters.  Since 7 and 12 will be saved as characters because of "5", the result will return `non-numeric argument to binary operator`.
    
    3. For the next two commands, either explain their results, or why they should produce errors. (3 points)

            z <- data.frame(z1="5",z2=7,z3=12)
            z[1,2] + z[1,3]
  The result of the above code is 19.  Unlike the previous question, this does not produce an error because z[1,2] and z[1,3] are not saved as characters given the structure of the data frame.
  
3. **Data structures** Give R expressions that return the following matrices and vectors (*i.e.* do not construct them manually). (3 points each, 12 total)

    1. $(1,2,3,4,5,6,7,8,7,6,5,4,3,2,1)$
```{r}
c(seq(1:8), rev(seq(1:7)))
```
    2. $(1,2,2,3,3,3,4,4,4,4,5,5,5,5,5)$
```{r}
rep(seq(1:5), seq(1:5))
```
    3. $\begin{pmatrix}
      0 & 1 & 1  \\
      1 & 0 & 1  \\
      1 & 1 & 0  \\
    \end{pmatrix}$
```{r}
m <-matrix(0,3,3)
m[lower.tri(m)] <- 1
m[upper.tri(m)] <- 1
m
```
    4. $\begin{pmatrix}
      1 & 2 & 3 & 4 \\
      1 & 4 & 9 & 16 \\
      1 & 8 & 27 & 64  \\
      1 & 16 & 81 & 256 \\
      1 & 32 & 243 & 1024  \\
    \end{pmatrix}$
```{r}
mat1 <-rep(seq(1:4),5)
mat2 <- matrix(mat1,nrow=5,ncol=4,byrow=TRUE)
mat3 <- seq(1:5)
mat4 <- mat2**mat3
mat4
```
4. **Basic programming** (10 points)

    1. Let $h(x,n)=1+x+x^2+\ldots+x^n = \sum_{i=0}^n x^i$. Write an R program to calculate $h(x,n)$ using a `for` loop. As an example, use `x = 5` and `n = 2`. (5 points)
```{r}
sum_h <- function(x, n) {
  sum = 0
  for(i in 0:n) {
    sum = sum + x**i
  }
  print(sum)
}
sum_h(5,2)
```
    1. If we list all the natural numbers below 10 that are multiples of 3 or 5, we get 3, 5, 6 and 9. The sum of these multiples is 23. Write an R program to perform the following calculations. (5 points)

        1. Find the sum of all the multiples of 3 or 5 below 1,000. (3, [euler1])
```{r}
l = (0)
for (i in 1:999){
  if (i%%3 == 0 || i%%5 == 0){
    l = append(l, i)
  }
}
sum(l)

three <- floor(999/3)
five <- floor(999/5)
both <- floor(999/15)
3*three*(three+1)/2 + 5*five*(five+1)/2 - 15*both*(both+1)/2
```
        1. Find the sum of all the multiples of 4 or 7 below 1,000,000. (2)
```{r}
four <- floor(1000000/4)
seven <- floor(1000000/7)
both <- floor(1000000/28)
4*four*(four+1)/2 + 7*seven*(seven+1)/2 - 28*both*(both+1)/2
```

```{r}
l = (0)
for (i in 1:1000000){
  if (i%%4 == 0 | i%%7 == 0){
    l = append(l, i)
  }
}
sum(l)
```
    1. Each new term in the Fibonacci sequence is generated by adding the previous two terms. By starting with 1 and 2, the first 10 terms will be $(1, 2, 3, 5, 8, 13, 21, 34, 55, 89)$. Write an R program to calculate the sum of the first 15 even-valued terms. (5 bonus points, [euler2])
```{r}
sum_fib <- function(n) {
  fib <- rep(NA, n*3)
  if (n <= 0){print(0)}
  else{
    fib[1] <- 0
    fib[2] <- 1
    count = 0
    sum = 0
    i = 3
    while(count < n){
      
      fib[i] = fib[i - 1] + fib[i-2]
      if(fib[i]%%2==0){
        sum = sum + fib[i]
        count = count + 1
      }
      i = i + 1
    }
  }
  print(sum)
}
sum_fib(15)
```
Some problems taken or inspired by projecteuler.

[euler1]: https://projecteuler.net/problem=1
[euler2]: https://projecteuler.net/problem=2
