---
title: 'Bios 6301: Assignment 5'
author: Lydia Yao
output: pdf_document
---

*Due Thursday, 14 October, 1:00 PM*

$5^{n=day}$ points taken off for each day late.

40 points total.

Submit a single knitr file (named `homework5.rmd`), along with a valid PDF output file. Inside the file, clearly indicate which parts of your responses go with which problems (you may use the original homework document as a template). Add your name as `author` to the file's metadata section. Raw R code/output or word processor files are not acceptable.

Failure to name file `homework5.rmd` or include author name may result in 5 points taken off.

### Question 1 ###

**15 points**

A problem with the Newton-Raphson algorithm is that it needs the derivative $f'$. If the derivative is hard to compute or does not exist, then we can use the *secant method*, which only requires that the function $f$ is continuous.

Like the Newton-Raphson method, the **secant method** is based on a linear approximation to the function $f$. Suppose that $f$ has a root at $a$. For this method we assume that we have *two* current guesses, $x_0$ and $x_1$, for the value of $a$. We will think of $x_0$ as an older guess and we want to replace the pair $x_0$, $x_1$ by the pair $x_1$, $x_2$, where $x_2$ is a new guess.

To find a good new guess x2 we first draw the straight line from $(x_0,f(x_0))$ to $(x_1,f(x_1))$, which is called a secant of the curve $y = f(x)$. Like the tangent, the secant is a linear approximation of the behavior of $y = f(x)$, in the region of the points $x_0$ and $x_1$. As the new guess we will use the x-coordinate $x_2$ of the point at which the secant crosses the x-axis.

The general form of the recurrence equation for the secant method is:

$$x_{i+1} = x_i - f(x_i)\frac{x_i - x_{i-1}}{f(x_i) - f(x_{i-1})}$$

Notice that we no longer need to know $f'$ but in return we have to provide *two* initial points, $x_0$ and $x_1$.

**Write a function that implements the secant algorithm.** Validate your program by finding the root of the function $f(x) = \cos(x) - x$. Compare its performance with the Newton-Raphson method -- which is faster, and by how much?  For this example $f'(x) = -\sin(x) - 1$.

```{r}
start_time <- Sys.time()
fofx <- function(x) {cos(x) - x}
secant <- function(x1,x2) {
  x2 -fofx(x2)*(x2 - x1)/(fofx(x2) - fofx(x1))
}

error = 1e-7
y = Inf 
x1 = 0
x2 = 1
count = 0
while(y > error){
  xn = secant(x1,x2)
  x1 = x2
  x2 = xn
  y = fofx(xn)
  count = count + 1
}
end_time <- Sys.time()
print(end_time - start_time)
print(paste("Secant method took" ,count,"iterations starting at 0 and 1"))
```

```{r}
start_time <- Sys.time()
fofx_d <- function(x) {-sin(x) - 1}
NR <- function(x) {x - fofx(x)/fofx_d(x)}
x1 = 1
y = Inf
count = 0
while(abs(y) > error){
  xn = NR(x1)
  x1 = xn
  y = fofx(xn)
  count = count + 1
}
end_time <- Sys.time()
print(end_time - start_time)
print(paste("Newton-Raphson method took" ,count,"iterations starting at 1"))
```
  ```
  From the simulation above, we see that the Newton-Raphson method is indeed faster by
  around 0.0112 seconds.  
  ```
### Question 2 ###

**20 points**

The game of craps is played as follows (this is simplified). First, you roll two six-sided dice; let x be the sum of the dice on the first roll. If x = 7 or 11 you win, otherwise you keep rolling until either you get x again, in which case you also win, or until you get a 7 or 11, in which case you lose.

Write a program to simulate a game of craps. You can use the following snippet of code to simulate the roll of two (fair) dice:

```{r}
roll <- function() sum(ceiling(6*runif(2)))
```

1. The instructor should be able to easily import and run your program (function), and obtain output that clearly shows how the game progressed.  Set the RNG seed with `set.seed(100)` and show the output of three games. (lucky 13 points)
```{r}
set.seed(100)
craps <- function() {
  first <- 0
  count <- 1
  while(TRUE) {
    x <- roll()
    print(paste("Roll number", count, ":", x))
    if ((x == 7 | x == 11) && first == 0) {
      return("You Won!!")
    }
    if (first == 0) {first = x}
    if (x == 7 || x == 11) {
      return("You Lose.")
    }
    count = count + 1
  }
  
}
cat("Game 1\n")
craps()
cat("Game 2\n")
craps()
cat("Game 3\n")
craps()
```

1. Find a seed that will win ten straight games.  Consider adding an argument to your function that disables output. Show the output of the ten games. (7 points)
```{r eval = FALSE}
library(sjmisc)
seed = 63398
  #1111480
while(TRUE) {
  print(seed)
  set.seed(seed)
  result = rep(NA, 30)
  for(i in 1:30){
      x <- capture.output(craps())
      if(str_contains(x, "You Won!!")) { result[i] = 1}
      else{result[i]= 0}
  }
  result
  consec <- rle(result)
  if (any(consec$lengths>=10 & consec$values==1)) {
    print(result)
    break
  }
  seed = seed + 1
}
```

```{r eval=FALSE}
library(sjmisc)
seed = 1
set.seed(seed)
result = rep(NA, 30)
  for(i in 1:17){
      x <- capture.output(craps())
      print(x)
      cat("\n")
      if(str_contains(x, "You Won!!")) { result[i] = 1}
      else{result[i]= 0}
  }
```


### Question 3 ###

**5 points**

This code makes a list of all functions in the base package:

```{r}
objs <- mget(ls("package:base"), inherits = TRUE)
funs <- Filter(is.function, objs)
```

Using this list, write code to answer these questions.

1. Which function has the most arguments? (3 points)
```{r}
#l <- function(x) length(formals(x))

l <- sapply(funs, function(x) length(formals(x)))
slength <- sort(l, decreasing = TRUE)
slength[1]
```

1. How many functions have no arguments? (2 points)
```{r}
no_arguments <- l[l == 0]
length(no_arguments)
```

Hint: find a function that returns the arguments for a given function.
