---
title: 'Bios 6301: Assignment 8'
author: 'Lydia Yao'
output: pdf_document
---

*Due Tuesday, 16 November, 1:00 PM*

$5^{n=day}$ points taken off for each day late.

30 points total.

Submit a single knitr file (named `homework8.rmd`), along with a valid PDF output file. Inside the file, clearly indicate which parts of your responses go with which problems (you may use the original homework document as a template). Add your name as `author` to the file's metadata section. Raw R code/output or word processor files are not acceptable.

Failure to name file `homework8.rmd` or include author name may result in 5 points taken off.

### Question 1 ###

**15 points**

Install the `readxl` package and run the following

```{r}
library(readxl)
fn <- 'icd10.xlsx'
if(file.access(fn, mode = 4) == -1) {
    url <- "https://www.cdc.gov/nhsn/xls/icd10-pcs-pcm-nhsn-opc.xlsx"
    download.file(url, destfile = fn, mode = 'wb')
}
dat <- readxl::read_excel(fn, sheet = 2)
```

1. Show the class of `dat`. (1 point)
```{r}
class(dat)
```

2. Show the methods available for objects of the given class (if there are multiple classes, show methods for all classes). (3 points)
```{r}
methods(class = 'data.frame')
methods(class = 'tbl')
methods(class = 'tbl_df')
```

3. If you call `print(dat)`, what print method is being dispatched? (1 point)
  ```
  The print method will dispatch the print under tbl_df.
  ```

4. Set the class of `dat` to be a data.frame. (1 point)
```{r}
class(dat) <- c('data.frame')
```

5. If you call `print(dat)` again, what print method is being dispatched? (1 point)
  ```
  This time, the print method will dispatch the print under data.frame.
  ```

Define a new generic function `nUnique` with the code below.

```{r}
nUnique <- function(x) {
    UseMethod('nUnique')
}
```

6. Write a default method for `nUnique` to count the number of unique values in an element. (2 points)
```{r}
nUnique <- function(x) {
  length(unique(x))
}
```
7. Check your function (2 points)

```{r, eval = FALSE}
nUnique(letters) # should return 26
nUnique(sample(10, 100, replace = TRUE)) # should return 10 (probably)
```

8. Write a data.frame method for `nUnique` to operate on data.frame objects.
This version should return counts for each column in a data.frame. (2 points)
```{r}
nUnique.data.frame <- function(x) {
  output <- matrix(NA, nrow = 1, ncol = ncol(x))
  for(i in seq(ncol(x))){
    output[,i] <- length(unique(x[,i]))
  }
  colnames(output) <- names(x)
  output
}
```
9. Check your function (2 points)

```{r, eval = FALSE}
nUnique.data.frame(dat)
```

### Question 2 ###

**15 points**

Programming with classes.  The following function will generate random patient information.

```{r}
makePatient <- function() {
  vowel <- grep("[aeiou]", letters)
  cons <- grep("[^aeiou]", letters)
  name <- paste(sample(LETTERS[cons], 1), sample(letters[vowel], 1), sample(letters[cons], 1), sep='')
  gender <- factor(sample(0:1, 1), levels=0:1, labels=c('female','male'))
  dob <- as.Date(sample(7500, 1), origin="1970-01-01")
  n <- sample(6, 1)
  doa <- as.Date(sample(1500, n), origin="2010-01-01")
  pulse <- round(rnorm(n, 80, 10))
  temp <- round(rnorm(n, 98.4, 0.3), 2)
  fluid <- round(runif(n), 2)
  list(name, gender, dob, doa, pulse, temp, fluid)
}
```

1. Create an S3 class `medicalRecord` for objects that are a list with the named elements `name`, `gender`, `date_of_birth`, `date_of_admission`, `pulse`, `temperature`, `fluid_intake`. Note that an individual patient may have multiple measurements for some measurements.  Set the RNG seed to `8` and create a medical record by taking the output of `makePatient`.  Print the medical record, and print the class of the medical record. (5 points)
```{r}
set.seed(8)
j <- makePatient()
class(j) <- 'medicalRecord'
print(j)

```
2. Write a `medicalRecord` method for the generic function `mean`, which returns averages for pulse, temperature and fluids. Also write a `medicalRecord` method for `print`, which employs some nice formatting, perhaps arranging measurements by date, and `plot`, that generates a composite plot of measurements over time.  Call each function for the medical record created in part 1. (5 points)
```{r}
mean.medicalRecord <- function(x){
  output <- matrix(NA, nrow = 1, ncol = 3)
  x <- x[c(5,6,7)]
  for(i in 1:3){
    output[,i] <- sapply(x[i], mean, na.rm = TRUE)
  }
  colnames(output) <- c("pulse", "temperature", "fluids")
  output
}

print.medicalRecord <- function(x){
  if(x[2] == 2) {gender = "male"}
  else{gender = "female"}
  
  dob = as.Date(as.integer(as.character.POSIXt(x[3])), origin="2010-01-01")
  
  four = sapply(x[4], as.character.POSIXt)
  four = sort(four)
  four = paste(four, collapse = ", ")
  
  cat(sprintf("name: %s\ngender: %s\ndob: %s\ndoa: %s\npulse: %s\ntemp: %s\nfluid: %s", 
              x[1], gender, dob, four, substr(x[5],3,nchar(x[5]) - 1), substr(x[6],3,nchar(x[6]) - 1), substr(x[7],3,nchar(x[7]) - 1)),"\n")
}
library(ggplot2)
plot.medicalRecord <- function(x){
  
  dates = sapply(x[4], as.character.POSIXt)
  dates = sapply(dates, as.Date)
  data <- data.frame(
    doa = as.Date(dates, origin="2010-01-01"),
    pulse = x[5],
    temp = x[6],
    fluid = x[7]
  )
  colnames(data) <- c("doa", "pulse", "temp", "fluid")
  ggplot(data) + geom_line(aes(x = doa, y = pulse, color = "black")) + geom_line(aes(x = doa, y = temp,color = "blue")) + geom_line(aes(x = doa, y = fluid, color = "red")) + scale_color_identity(name = "Variable",
                          breaks = c("black", "blue", "red"),
                          labels = c("pulse", "temp", "fluid"),
                          guide = "legend") + xlab("Date") + ylab("Number")
}
mean(j)
print(j)
plot(j)
```

3. Create a further class for a cohort (group) of patients, and write methods for `mean` and `print` which, when applied to a cohort, apply mean or print to each patient contained in the cohort. Hint: think of this as a "container" for patients.  Reset the RNG seed to 8 and create a cohort of ten patients, then show the output for `mean` and `print`. (5 points)
```{r}
set.seed(8)
cohort <- function(n){
  result = makePatient()
  for(i in 1:(n-1)){
    result = c(result, makePatient())
  }
  result
}
j2 <- cohort(10)
class(j2) <- 'cohortRecord'

mean.cohortRecord <- function(x){
  output <- matrix(NA, nrow = length(x)/7, ncol = 3)
  pulse = x[seq(5, length(x), 7)]
  temp = x[seq(6, length(x), 7)]
  fluids = x[seq(7, length(x), 7)]
  for(i in 1:as.integer(length(x)/7)){
    output[i,1] <- sapply(pulse[i], mean, na.rm = TRUE)
    output[i,2] <- sapply(temp[i], mean, na.rm = TRUE)
    output[i,3] <- sapply(fluids[i], mean, na.rm = TRUE)
  }
  colnames(output) <- c("pulse", "temperature", "fluids")
  output
}

print.cohortRecord <- function(x){
  count = 1
  for(i in 1:length(x)){
    if(count ==1){
      print(paste("name: ", x[i]))
      count = count + 1
    } else if (count ==2) {
      if(x[i] == 2) {print("gender: male ")}
      else{print("gender: female ")}
      count = count + 1
    } else if (count ==3){
      dob = as.Date(as.integer(as.character.POSIXt(x[i])), origin="2010-01-01")
      print(paste("dob: ", dob))
      count = count + 1
    } else if (count ==4){
      four = sapply(x[i], as.character.POSIXt)
      four = sort(four)
      four = paste(four, collapse = ", ")
      print(paste("doa: ", four))
      count = count + 1
    } else if (count ==5){
      print(paste("pulse: ", substr(x[i],3,nchar(x[i]) - 1)))
      count = count + 1
    } else if (count ==6){
      print(paste("temp: ", substr(x[i],3,nchar(x[i]) - 1)))
      count = count + 1
    } else if (count ==7){
      print(paste("fluid: ", substr(x[i],3,nchar(x[i]) - 1)))
      count = 1
      cat("\n")
    }
  }
}
mean.cohortRecord(j2)
print.cohortRecord(j2)
```



